% Exercise 2.3
\section{}

Which states are superpositions with respect to the standard basis, and which are not?
For each state that is a superposition, give a basis with respect to which it is not a
superposition.

\begin{multicols}{2}
\begin{enumerate}[label=\alph*.,ref={Ex.~\thesection\alph*}]
    \label{ex:2p3}
        % a.
        \item $\ket{+}$
        % b.
        \item $\spket{}{+}{+}{-}$
        % c.
        \item $\spket{}{+}{-}{-}$
        % d.
        \item $\tfrac{\sqrt{3}}{2} \ket{+} - \tfrac{1}{2} \ket{-}$
        % e.
        \item $\spket{}{\ibf}{-}{- \ibf}$
        % f.
        \item $\spket{}{0}{-}{1}$
    \end{enumerate}
\end{multicols}

{\Sol}
\begin{enumerate}[label=\alph*.,ref={Sol.~\thesection\alph*}]
\label{sol:2p3}
    % a.
    \item Superposition of the standard basis; not a superposition of the Hadamard basis (trivial).
    % b.
    \item Recall from \ref{sol:2p2g} that this state is just $\ket{0}$.
    Therefore, it is not a superposition of the standard basis.
    % c.
    \item This state is just $\ket{1}$; not a superposition of the standard basis.
    % d.
    \item \label{sol:2p2d} Maybe this is not too clear at first glance -- let's expand the state:
    \begin{align*}
        \tfrac{\sqrt{3}}{2} \ket{+} - \tfrac{1}{2} \ket{-} &= \tfrac{\sqrt{3}}{2} \left (\spket{}{0}{+}{1} \right )  - \tfrac{1}{2} \left (\spket{}{0}{-}{1} \right ) \\
        &= \tfrac{1}{2\sqrt{2}} \left ( ({\sqrt{3} - 1}) \ket{0} + ({\sqrt{3} + 1}) \ket{1} \right )
    \end{align*}
    Indeed, this is a superposition of the standard basis. We can easily construct a basis in which this state is
    not a superposition by choosing one of the bases as the state itself, and then finding an orthogonal state to it.
    The orthogonal state can be constructed trivially by negating the sign of one of $\ket{-}$ and swapping the amplitudes,
    \begin{align*}
        \tfrac{\sqrt{3}}{2} \ket{+} - \tfrac{1}{2} \ket{-} \to \tfrac{1}{2} \ket{+} + \tfrac{\sqrt{3}}{2} \ket{-}.
    \end{align*}
    So, a basis in which this state is not a superposition is,
    \begin{align*}
        \left \{ \tfrac{\sqrt{3}}{2} \ket{+} - \tfrac{1}{2} \ket{-}, \tfrac{1}{2} \ket{+} + \tfrac{\sqrt{3}}{2} \ket{-} \right \}.
    \end{align*}
    Note that this ``swap and negate'' trick does not always work; see \ref{sol:2p7d}.
    % e.
    \item We know from \ref{sol:2p2h} that this state is $\ibf \ket{1}$, so it is not a superposition of the standard basis.
    % f.
    \item This is clearly a superposition of the standard basis, but not of the Hadamard basis.
\end{enumerate}
